Key Derivation From Noisy Sources With More Errors Than Entropy

### 2014-09-15 trunk/constantRate/ConstCompFuzzExt-reorg.tex
working on intro

Constructions of fuzzy extractors are limited by the tension between security and correctness guarantees: if we allow for higher error tolerance $t$, then we also need higher starting entropy $m$. The reason for this tension is simple: if an adversary who knows $p$ can guess any $w'$ within distance $t$ of $w$, it will be able to easily obtain the true $r$ by running $\rep$.
%If $t$ is larger, then a single guess for $w'$ may be within distance $t$ of more $w$ values, thus increasing adversarial probability of winning. 
In fact, if $t$ is high enough that there are $2^m$ points in a ball of radius $t$, then there exists a distribution of $w$ of min-entropy $m$  \emph{contained entirely in a single ball}.  For this distribution, an adversary can run $\rep$ on the center of this ball and always learn the key $r$.

More generally, let $B_t$ denote the number of points in a ball of radius $t$. For any $m$ and $t$, there is a distribution of min-entropy $m$ such that the adversary can guess a correct $w'$ with probability $1/\lceil( 2^m/B_t) \rceil\approx B_t 2^{-m}$: the distribution consists of the uniform distribution over all points in several non-overlapping balls of radius $t$ (the metric space must be large enough for these balls not to intersect). We thus call $m-\log B_t$ the \emph{minimum usable} entropy, denoted by $\Huse$. The previous paragraph shows that  no fuzzy extractor can handle all distributions of a given min-entropy $m$ if  $\Huse\le 0$.

Prime candidate sources for authentication have $\Huse\le 0$.  As an example, the iris is believed to be the best biometric for high security applications~\cite{prabhakar2003biometric}.  Daugman estimates irises contain $249$ bits of entropy~\cite{daugman2004}.  Daugman uses specialized wavelets to derive a $2048$ bit string called an iris code.  Let the outcome of this transform~(on different irises) define a distribution $w$.  The precise number of errors that must be tolerated depends on the desired false reject rate (how often the correct key is produced).  For a false reject rate of $\le 80\%$, a $t$ of approximately $205$ is required.  We have the following calculation for $\Huse$:
\[
\Huse = \Hoo(W) - \log |B_t|
= 249 - \log \sum_{i=0}^{205} {2048 \choose i} \approx -707.
\]
There is considerable subsequent research~\cite{gentile2009slic,gentile2009efficient,rathgeb2011combining} to Daugman but it does not affect $\Huse$ dramatically.%\footnote{The work of Hao et al.~\cite[Section 4.3]{hao2006combining} provides a similar analysis but their calculation underestimates the number of possible error patterns; their calculation, which we cannot confirm, is $\Huse \approx 44$.}

\paragraph{Our Contributions}
We provide the first constructions of fuzzy extractors that can be used for a large class of distributions with $\Huse\le 0$ over the Hamming metric on $\mathcal{Z}^\ell$ for an alphabet $\mathcal{Z}$.

\bnote{rewrite this!}
Our first construction is secure when symbols in $w$
each have individual super-logarithmic min-entropy, even if they are arbitrarily correlated. Moreover,
a constant fraction of symbols in $w$ may have little entropy, as long as knowledge of their values does not reduce the entropy of the high-entropy symbols too much (see \defref{def:block guessable}).  This construction can correct a constant fraction of errors~($\Theta(\ell)$).  However, this construction requires a large alphabet size~(super-polynomial in the security parameter).

We improve the entropy and alphabet size requirements in the second construction, which requires only a constant fraction of the symbols $w$ to have constant min-entropy conditioned on the previous symbols.  Our second construction supports an arbitrary alphabet size~(though it only achieves $\Huse\le 0$ when the alphabet size is super-constant).  
However, these improvements come at a price to error-tolerance: it tolerates a sub-constant fraction of errors~($o(\ell)$).

\consref{cons:sampling} comes with an additional feature: reusability.  Boyen~\cite{Boyen2004} defined reusable fuzzy extractors: multiple readings $w_1,..., w_q$ are enrolled and the adversary sees the corresponding public values $p_1,..., p_q$ and possibly some of the derived keys.  The enrolled values $w_i$ and $w_j$ are not assumed to be equal, and so information leakage may add up.  Indeed, Boyen showed that with an unbounded number of reuses it is only possible to support limited correlations between $w_i$ and $w_j$~\cite[Theorem 11]{Boyen2004}.  Our second construction allows for arbitrary correlation between $w_i$ and $w_j$ against computational adversaries\footnote{Our restriction to computational adversaries implies that the adversary only sees a polynomially bounded number of reuses.  It remains an open problem to construct a polynomially reusable fuzzy extractor against unbounded adversaries.} as long as each marginal $W_i$ is an admissible distribution.  %We do not considering the joint distribution $W_1,..., W_q$.

>>>>

Constructions of fuzzy extractors are limited by the tension between security and correctness guarantees: if we allow for higher error tolerance $t$, then we also need higher starting entropy $m$. The reason for this tension is simple: if an adversary who knows $p$ can guess any $w'$ within distance $t$ of $w$, then it will be able to easily obtain the true $r$ by running $\rep$.
%If $t$ is larger, then a single guess for $w'$ may be within distance $t$ of more $w$ values, thus increasing adversarial probability of winning. 
In fact, if $t$ is high enough that there are $2^m$ points in a ball of radius $t$, then there exists a distribution of $w$ of min-entropy $m$  \emph{contained entirely in a single ball}.  For this distribution, an adversary can run $\rep$ on the center of this ball and always learn the key $r$.
Thus, if the security guarantee of a given fuzzy extractor holds for \emph{any} source of a given min-entropy $m$ and the correctness guarantees holds for any $t$ errors, then $m$ must   be greater than $\log B_t$, where $B_t$ denotes the number of points in a ball of radius $t$.  If a source fails this condition, we will says that it has \emph{more errors than entropy}.

%More generally, for any $m$ and $t$, there is a distribution of min-entropy $m$ such that the adversary can guess a correct $w'$ with probability $1/\lceil( 2^m/B_t) \rceil\approx B_t 2^{-m}$: the distribution consists of the uniform distribution over all points in several non-overlapping balls of radius $t$ (the metric space must be large enough for these balls not to intersect). We thus call $m-\log B_t$ the \emph{minimum usable} entropy, denoted by $\Huse$. The previous paragraph shows that  no fuzzy extractor can handle all distributions of a given min-entropy $m$ if  $\Huse\le 0$.

%Prime candidate sources for authentication have $\Huse\le 0$.  
Unfortunately, prime candidate sources for authentication have more errors than entropy. 
As an example, the iris is believed to be the best biometric for high security applications~\cite{prabhakar2003biometric}.  Daugman~\cite{daugman2004} designs a transform called IrisCode (using specialized wavelets) to derive a $2048$ bit string from an iris.  Let the outcome of this transform (on different irises) define a distribution $w$.  The precise number of errors that must be tolerated depends on the desired false reject rate.  Even allowing for a false reject rate of up to $80\%$ (i.e., permitting correctness to fail $80\%$ of the time), a $t$ of approximately $205$ is required. Thus,
\[
\log |B_t|
= \log \sum_{i=0}^{205} {2048 \choose i} \approx 956\,.
\]
In contrast, $w$ is estimated to have about $249$ bits of entropy~\cite{daugman2004}.
There is considerable subsequent research~\cite{gentile2009slic,gentile2009efficient,rathgeb2011combining}, but it does not affect the above comparison dramatically.%\footnote{The work of Hao et al.~\cite[Section 4.3]{hao2006combining} provides a similar analysis but their calculation underestimates the number of possible error patterns; their calculation, which we cannot confirm, is $\Huse \approx 44$.}

The situation with reusability is even worse: the only known construction of reusable fuzzy extractors \cite{Boyen2004} requires very particular relationships between $w_i$ values, which are unlikely to hold in any practical source.

\paragraph{Our Contributions: Handling Sources with More Errors than Entropy}
We provide the first constructions of fuzzy extractors that can be used for large classes of sources that have more errors than entropy.  Our constructions work for Hamming errors for strings $w$ of length $\ell$ over some alphabet $\mathcal{Z}$. Naturally, as argued above, these constructions cannot work for all sources of a given entropy; each construction comes with a constraint on the sources for which it is secure.

Our first construction provides information-theoretic security.  It can correct a constant fraction of errors, but requires that a constant fraction of the symbols contribute fresh entropy, even conditioned on previous symbols. It requires an alphabet size that is super constant in the security parameter. \lnote{keep this last sentence?}

We then switch to computational security to obtain constructions with additional features. Our second construction, like the first, also requires that a constant fraction of symbols contribute fresh entropy. However, it provides reusability. The reusability we obtain is very strong:  security holds even if the multiple readings $w_i$ used in $\gen$ are \emph{arbitrarily correlated}, as long as each $w_i$ \emph{individually} satisfies the above conditions as a distribution.
We note that such strong reusability was shown to be impossible in information-theoretically secure fuzzy extractors~\cite[Theorem 11]{Boyen2004}.\lnote{really? I can't parse his theorem clearly enough} This construction, however, requires that the fraction of errors is subconstant.

Our third construction removes the need for fresh entropy in the symbols. 
It is secure when symbols in $w$
each have individual super-logarithmic min-entropy, even if they are arbitrarily correlated. Moreover,
a constant fraction of symbols in $w$ may have little entropy, as long as knowledge of their values does not reduce the entropy of the high-entropy symbols too much (see \defref{def:block guessable}).  This construction can correct a constant fraction of errors.  However, this construction requires a large alphabet size~(super-polynomial in the security parameter).

%\bnote{rewrite this!}
%Our first construction is secure when symbols in $w$ each have individual super-logarithmic min-entropy, even if they are arbitrarily correlated. Moreover, a constant fraction of symbols in $w$ may have little entropy, as long as knowledge of their values does not reduce the entropy of the high-entropy symbols too much (see \defref{def:block guessable}).  This construction can correct a constant fraction of errors~($\Theta(\ell)$).  However, this construction requires a large alphabet size~(super-polynomial in the security parameter).
%
%We improve the entropy and alphabet size requirements in the second construction, which requires only a constant fraction of the symbols $w$ to have constant min-entropy conditioned on the previous symbols.  Our second construction supports an arbitrary alphabet size~(though it only achieves $\Huse\le 0$ when the alphabet size is super-constant).  However, these improvements come at a price to error-tolerance: it tolerates a sub-constant fraction of errors~($o(\ell)$).  
%
%\consref{cons:sampling} comes with an additional feature: reusability.  Boyen~\cite{Boyen2004} defined reusable fuzzy extractors: multiple readings $w_1,..., w_q$ are enrolled and the adversary sees the corresponding public values $p_1,..., p_q$ and possibly some of the derived keys.  The enrolled values $w_i$ and $w_j$ are not assumed to be equal, and so information leakage may add up.  Indeed, Boyen showed that with an unbounded number of reuses it is only possible to support limited correlations between $w_i$ and $w_j$~\cite[Theorem 11]{Boyen2004}.  Our second construction allows for arbitrary correlation between $w_i$ and $w_j$ against computational adversaries\footnote{Our restriction to computational adversaries implies that the adversary only sees a polynomially bounded number of reuses.  It remains an open problem to construct a polynomially reusable fuzzy extractor against unbounded adversaries.} as long as each marginal $W_i$ is an admissible distribution.  %We do not considering the joint distribution $W_1,..., W_q$.
